Curriculum and Transfer Learning of Hierarchical Locomotion Policies

### 2018-05-31 paper/main.tex
Write introduction, aspirational experiments introduction, and some ideas for policy architecture

\newcommand{\highlight}[1]{{\leavevmode\color{blue} #1}}
\newcommand{\warning}[1]{{\leavevmode\color{red}Warning: #1}}

@@

@@

  \item \cite{peng2017learning} uses bumpy ground to test robustness of learned policies, but again the bumps are not actually perceived and cannot be planned around. The bump height thus must remain small relative to the agent. \highlight{We should compare this quantitatively, later.} Also, this is 2D.
  \item \cite{manchester2011stable} This paper stands out in that they derive a provably stable controller that they test on a physical robot. \highlight{I haven't read it in detail yet, though.}
\end{itemize}

@@

  \item \cite{hausman2018learning} is an interesting theoretical paper showing how to get diverse meta-controls without imposing much structure on the interface between the HLC and the LLC.
    However, in our case, we have a prior idea about a good hierarchical breakdown of the tasks (namely, a footstep plan). \highlight{I had some trouble with this paper; I should probably reread it and update this summary.}

@@

\begin{itemize}
  \item At the lowest level, we will probably be using some kind of stable PD controller, in order to make the reinforcement learning problem as easy as possible. \highlight{I need to read these papers:} \cite{tan2011stable}, \cite{shrivastava2013stable}
  \item \cite{peng2015terrain} introduces a very interesting way of parametrizing the action space in terms of reference points for PD controllers, as well as ``virtual forces''. We will want to do something similar.
  \item We might want to use something simple, like a linear controller as proposed by \cite{mania2018simple} to see if that works well.
    It seems to do a good job on OpenAI benchmarks, at least.
  \item Otherwise, we'll probably end up using TRPO or PPO or something (get citation later\dots)
  \item Other things to consider: prioritized experience replay, hindsight experience replay, universal value function approximators.
\end{itemize}

@@

\item \textbf{Introduction:} Robotic locomotion on uneven terrain is an important problem. \highlight{Visak, I would be interested in your high-level reasons for being interested in this problem. I'm not sure I can motivate the problem as well as you can.} Previous approaches have not solved this problem. For instance, in \href{https://www.youtube.com/watch?v=hx_bgoTF7bs&feature=youtu.be&t=1m28s}{this video} the agent stumbles on rubble, and DeepLoco only walks on flat terrain. \highlight{More examples would help, to make sure that this problem is interesting.} We hope to achieve efficient locomotion across bumpy ground using a combination of three techniques.
  \begin{enumerate}
    \item Curriculum learning: Train the agent first on simpler environments, like flat ground and gradual slopes, and gradually increase the difficulty.
    \item Hierarchical controllers: We use a hierarchical architecture where the HLC plans foot placement and the LLC executes the individual steps. We hypothesize that this architecture will more easily handle the discontinuities caused by contact dynamics. We also hypothesize that, with this architecture, it will be easier to transfer knowledge between tasks in the curriculum.
    \item Model-based planning: To help with sample efficiency, the HLC has access to a physics model for the environment. \highlight{I'm not sure how to accomplish this. Physics models for legged robots are probably hard, especially when the ground is not flat. Maybe some kind of \emph{learned} model would work.}
  \end{enumerate}

@@

\highlight{This section could probably use more detail; suggestions are welcome.
For example, do we need to modify this architecture to better handle contact dynamics?
Is there some way we can incorporate a model to help the HLC make better plans?}

@@

\item \textbf{Why this might not be a good idea:}
  \highlight{Why would *hierarchical* RL be a particularly good way to do transfer learning for this series of tasks?}
  We will need to implement several other baselines to compare against.
  \highlight{What should those baselines be? (What other approaches have people used for locomotion on bumpy ground?)
Is the ``bumpy ground'' task too difficult? (Byron thinks so.)
Will the simulated robot need toes?}
I haven't worked with locomotion before, so I don't have a good sense for this.

>>>>

\newcommand{\nhatch}[1]{{\leavevmode\color{blue} Nathan: #1}}

@@

\section{Introduction}

Locomotion is hard because the changing contacts make the dynamics difficult to model.
Recent deep reinforcement learning approaches try to learn this model implicitly via the policy network, which has trouble learning such complicated models, possibly due to the limitations of current optimization methods (like policy gradient).
Such approaches fail in environments that require predictive planning at the level of individual contacts, such as running across irregular terrain, where small miscalculations have disproportionally large effects.

We aim to simplify the learning problem by imposing prior structure on the policy: we split it into a two-level hierarchy.
The higher level breaks the plan into smaller chunks, which are easier to achieve.
The lower level must then execute these smaller chunks.
Each chunk is a single footstep.
This choice simplifies the dynamics that the lower level needs to model, because each step involves no (or very few) changes of contact.
It also simplifies the upper level learning problem, since the controller does not need to handle details of torque, other than knowing which step plans can actually be achieved.

\section{Experiments}

Our experiments are designed to test two hypotheses.

Our primary hypothesis is that this hierarchical architecture can navigate a variety of complicated locomotion environments, learning faster and achieving better results than other reinforcement learning methods.
These locomotion environments include flat ground, inclines, stairs, random inclines, stepping stones, hurdles, and combinations of the above.

Our two metrics for ``learning faster'' are (1) number of simulation frames of experience collected, and (2) total computation time.
We consider both necessary, because the relative expense of more experience versus more computation may vary across application domains.

We collect both quantitative and qualitative metrics to measure ``achieving better results''.
Quantitatively, we compare the average rate of forward progress.
We also consider the robustness of the model to force perturbations and various measures of the ``difficulty'' of the terrain (such as stair height).
Qualitatively, we document motion quality by means of supplementary videos, which we encourage the reader to consult.

Our secondary hypothesis is that the learned policy can be run in real-time while still achieving results comparable to more expensive, model-based optimization methods.

This hypothesis is difficult to test, because model-based optimization methods are difficult to implement for direct comparison.
Thus, our metric is simply the percent real time necessary to compute the control for the next time step, which we compare to numbers reported in optimization-based papers.

Additionally, we conduct ablative analysis of the importance of various components of our approach.
\nhatch{In particular, we want to consider various different HLC / LLC interfaces that we could have used.
For example: should we include a target foot height that the LLC should reach?
(I think this would help when going over hurdles.)
For another example (assuming we use some kind of PD control at the lowest level): how much does performance degrade if we don't use PD control?
(That is, if the LLC outputs raw torques.)}

\section{Policy architecture}

\nhatch{Ideally, maybe we should design this so that walking and running can both be achieved in the same episode.}

\nhatch{Should there be some ensemble of low-level policy networks that the HLC can choose between?
For example, we could take the policy with the highest estimated value given the target step location.
Having some discrete choices like this might be easier than using solely continuous parameters.}

\textbf{High-level observations.}
We assume that perception is solved.
The high-level controller receives proprioceptive information and environmental information.
The environmental information consists of a list of potential surfaces to step on, specified in terms of relative displacement from the agent and surface normal.
\nhatch{Maybe, for polygonal surfaces, we could also report the surface ``size'' (e.g. largest-radius circle that fits inside the polygon).}

\textbf{High-level actions.}
High-level actions are chosen on each heel-down event (once per footstep).
At these events, the HLC outputs a target for the next footstep that it expects the LLC can achieve.
This target consists of several parameters: \nhatch{TODO think about this some more}
\nhatch{It might be possible to incorporate some optimization / search / planning into the HLC, because its actions happen infrequently.
But this would require some objective to be optimized, which might involve implementing a model like a spring-loaded inverted pendulum (SLIP), or inventing some heuristic objective that might characterize good plans.
We might want to implement a model like SLIP anyway, in order to use it as an expert to train a neural network for the HLC (similar to AlphaGo Zero and ExIt).
In any case, only the first step of the plan would be revealed to the LLC.}

\textbf{Low-level observations.}
The low-level controller receives proprioceptive information and the target location for the next footstep.
Unlike the HLC, it does not observe the external environment.

\textbf{Low-level actions.}
On the one hand, we could have low-level actions being the raw torques.
On the other hand, this might make the learning problem harder, so we could specify some more abstract actions and use PD controllers or other analytical formulas to derive the actual torques.
For instance, the output could be a target pose or end-effector position which is chosen more infrequently than the simulation framerate.
Then we could solve an inverse kinematics or inverse dynamics problem to calculate a target pose, which is then achieved using PD controllers.
\nhatch{We probably don't need all of this machinery.
For the initial proof of concept, it might make sense to see how we can do using just raw torques.}
\nhatch{TODO think about this some more.}

\textbf{PD control.}
\nhatch{We might have to rethink this abstraction layer once the ground includes inclines.
For instance, the PD controller might now have to handle different target ankle angles.}

\section{Timeline of work}

The goal is to submit this to ICRA on September 15.

\begin{itemize}
  \item Early June (2 weeks): Proof-of-concept: train a 2D walker low-level controller to follow an irregular step plan on flat ground, including some low-level PD control if necessary.
  \item Late June (2 weeks): Train low-level controller on stepping stones.
    Train high-level controller on flat ground, then on stepping stones.
    Compare performance on stepping stones to baseline deep RL approach (TRPO).
    If we can't get this to work (better than TRPO), we should consider abandoning the project.
  \item July (4 weeks): Implement and train on inclines, stairs, and hurdles.
    Perform ablative analyses.
  \item Early August (2 weeks): I (Nathan) will be on vacation.
  \item Late August and early September (4 weeks): Last-minute experiments.
    Write and revise the paper.
  \item September 15: ICRA deadline.
\end{itemize}

@@

  \item \cite{peng2017learning} uses bumpy ground to test robustness of learned policies, but again the bumps are not actually perceived and cannot be planned around.
    The bump height thus must remain small relative to the agent.
    \nhatch{We should compare this quantitatively, later.} Also, this is 2D.
  \item \cite{manchester2011stable} This paper stands out in that they derive a provably stable controller that they test on a physical robot.
    \nhatch{I haven't read it in detail yet, though.}
\end{itemize}

@@

  \item \cite{hausman2018learning} is an interesting theoretical paper showing how to get diverse meta-controls without imposing much structure on the interface between the HLC and the LLC.
    However, in our case, we have a prior idea about a good hierarchical breakdown of the tasks (namely, a footstep plan).
    \nhatch{I had some trouble with this paper; I should probably reread it and update this summary.}

@@

\begin{itemize}
  \item At the lowest level, we will probably be using some kind of stable PD controller, in order to make the reinforcement learning problem as easy as possible.
    \nhatch{I need to read these papers:} \cite{tan2011stable}, \cite{shrivastava2013stable}
  \item \cite{peng2015terrain} introduces a very interesting way of parametrizing the action space in terms of reference points for PD controllers, as well as ``virtual forces''.
    We will want to do something similar.
  \item We might want to use something simple, like a linear controller as proposed by \cite{mania2018simple} to see if that works well.
    It seems to do a good job on OpenAI benchmarks, at least.
  \item Otherwise, we'll probably end up using TRPO or PPO or something (get citation later\dots)
  \item Other things to consider: prioritized experience replay, hindsight experience replay, universal value function approximators.
\end{itemize}

@@

\item \textbf{Introduction:} Robotic locomotion on uneven terrain is an important problem.
  \nhatch{Visak, I would be interested in your high-level reasons for being interested in this problem.
  I'm not sure I can motivate the problem as well as you can.} Previous approaches have not solved this problem.
  For instance, in \href{https://www.youtube.com/watch?v=hx_bgoTF7bs&feature=youtu.be&t=1m28s}{this video} the agent stumbles on rubble, and DeepLoco only walks on flat terrain.
  \nhatch{More examples would help, to make sure that this problem is interesting.} We hope to achieve efficient locomotion across bumpy ground using a combination of three techniques.
  \begin{enumerate}
    \item Curriculum learning: Train the agent first on simpler environments, like flat ground and gradual slopes, and gradually increase the difficulty.
    \item Hierarchical controllers: We use a hierarchical architecture where the HLC plans foot placement and the LLC executes the individual steps.
      We hypothesize that this architecture will more easily handle the discontinuities caused by contact dynamics.
      We also hypothesize that, with this architecture, it will be easier to transfer knowledge between tasks in the curriculum.
    \item Model-based planning: To help with sample efficiency, the HLC has access to a physics model for the environment.
      \nhatch{I'm not sure how to accomplish this.
      Physics models for legged robots are probably hard, especially when the ground is not flat.
      Maybe some kind of \emph{learned} model would work.}
  \end{enumerate}

@@

\nhatch{This section could probably use more detail; suggestions are welcome.
For example, do we need to modify this architecture to better handle contact dynamics?
Is there some way we can incorporate a model to help the HLC make better plans?}

@@

\item \textbf{Why this might not be a good idea:}
  \nhatch{Why would *hierarchical* RL be a particularly good way to do transfer learning for this series of tasks?}
  We will need to implement several other baselines to compare against.
  \nhatch{What should those baselines be? (What other approaches have people used for locomotion on bumpy ground?)
Is the ``bumpy ground'' task too difficult? (Byron thinks so.)
Will the simulated robot need toes?}
I haven't worked with locomotion before, so I don't have a good sense for this.
